Vegetation regrowth/encroachment (gjengroing)

### 2023-08-29 DRAFT_functional_plant_indicators_natopen.Rmd
natopen markdown finished

unique(ANO.natopen$hovedoekosystem_punkt)
unique(ANO.natopen$hovedtype_rute)
unique(ANO.natopen$kartleggingsenhet_1m2)
ANO.natopen$hovedtype_rute <- as.factor(ANO.natopen$hovedtype_rute)
ANO.natopen$kartleggingsenhet_1m2 <- as.factor(ANO.natopen$kartleggingsenhet_1m2)
summary(ANO.natopen$hovedtype_rute)
summary(ANO.natopen$kartleggingsenhet_1m2)
# we have no reference for overall T2 registrations, these 24 observations won't be evaluated
ANO.natopen[ANO.natopen$kartleggingsenhet_1m2=="T2",] # were largely inaccessible and thus mapped from arial pictures

@@

```{r}
ggplot(res.natopen.GRUK[res.natopen.GRUK$fp_ind=="Nitrogen2",], aes(x=fremmedartsdekning, y=scaled_value)) +
  geom_point() +
  xlab("Alien species cover (%)") + ylab("Scaled Nitrogen2 value (GRUK data)")
```
And indeed, we can see that very high alien species cover predicts poor scores in the Nitrogen2 indicator.

@@

CSR-R, ANO:
```{r}
# summary(res.natopen.ANO$kartleggingsenhet_1m2)
# 15 NiN-types, of which T2, T13 & T18 as main types don't have a reference
# many T13-types (rasmark) have too few vascular plants in the reference lists for getting a reference distribution
# C1, C2 and C10 in the ANO data have none
#-> so, 9 NiN-Types to plot

x11()
par(mfrow=c(3,3))

for ( i in sort(unique(res.natopen.ANO$kartleggingsenhet_1m2)[-c(2,8,9,12,13,15)]) ) {

@@

    plot(density( as.matrix(natopen.ref.cov[['RR']][,i]) ,na.rm=T),
         xlim=c(0,1), ylim=c(0,10), type="l", main=i,xlab='CSR-R value')
    points(density(res.natopen.ANO[res.natopen.ANO$fp_ind=="RR1" & res.natopen.ANO$kartleggingsenhet_1m2==i,]$original,na.rm=T),
           type="l", col="red")
    points(res.natopen.ANO[res.natopen.ANO$fp_ind=="RR1" & res.natopen.ANO$kartleggingsenhet_1m2==i,]$original,
           rep(0,length(res.natopen.ANO[res.natopen.ANO$fp_ind=="RR1" & res.natopen.ANO$kartleggingsenhet_1m2==i,]$original)),
           col="red")

@@

}
legend("topright", legend=c("reference","field data"), pch=c(NULL,1), lty=1, col=c("black","red"))

@@

# only two NiN types in GRUK, T2-C-7 and T2-C-8
par(mfrow=c(1,2))

for ( i in unique(res.natopen.GRUK$Kartleggingsenhet) ) {

@@

    plot(density( as.matrix(natopen.ref.cov[['Nitrogen']][,i]) ,na.rm=T),
         xlim=c(1,9), ylim=c(0,2), type="l", main=i,xlab='Nitrogen value')
    points(density(res.natopen.GRUK[res.natopen.GRUK$fp_ind=="Nitrogen1" & res.natopen.GRUK$Kartleggingsenhet==i,]$original,na.rm=T),
           type="l", col="red")
    points(res.natopen.GRUK[res.natopen.GRUK$fp_ind=="Nitrogen1" & res.natopen.GRUK$Kartleggingsenhet==i,]$original,
           rep(0,length(res.natopen.GRUK[res.natopen.GRUK$fp_ind=="Nitrogen1" & res.natopen.GRUK$Kartleggingsenhet==i,]$original)),
           col="red")

@@

}
legend("topright", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"))

```
The figure shows that deviations mainly occur in V2-C1, which represents limestone-poor swamp forests. According to the functional signature from the plant community composition many of the occurrences of this ecosystem type are too acidic and may have too low availability of nitrogen.

>>>>

unique(ANO.natopen$hovedoekosystem_punkt)
unique(ANO.natopen$hovedtype_rute)
unique(ANO.natopen$kartleggingsenhet_1m2)
ANO.natopen$hovedtype_rute <- as.factor(ANO.natopen$hovedtype_rute)
ANO.natopen$kartleggingsenhet_1m2 <- as.factor(ANO.natopen$kartleggingsenhet_1m2)
summary(ANO.natopen$hovedtype_rute)
summary(ANO.natopen$kartleggingsenhet_1m2)
# we have no reference for overall T2 registrations, these 24 observations won't be evaluated
ANO.natopen[ANO.natopen$kartleggingsenhet_1m2=="T2",] # were largely inaccessible and thus mapped from aerial pictures

@@

```{r}
colnames(res.natopen.GRUK)[38] <- "fremmedartsdekning"
ggplot(res.natopen.GRUK[res.natopen.GRUK$fp_ind=="Nitrogen2",], aes(x=fremmedartsdekning, y=scaled_value)) +
  geom_point() +
  xlab("Alien species cover (%)") + ylab("Scaled Nitrogen2 value (GRUK data)")
```
And indeed, we can see that very high alien species cover predicts almost only poor scores in the Nitrogen2 indicator.

@@

CSR-R, ANO:
```{r}
#summary(res.natopen.ANO[!is.na(res.natopen.ANO$original),]$kartleggingsenhet_1m2)
#sort(unique(res.natopen.ANO$kartleggingsenhet_1m2))
# 8 NiN-types with data to plot

par(mfrow=c(2,4))
for ( i in sort(unique(res.natopen.ANO$kartleggingsenhet_1m2))[c(1,6,7,10,12,13,14,15)] ) {

@@

    plot(density( as.matrix(natopen.ref.cov[['RR']][,i]) ,na.rm=T),
         xlim=c(0,1), ylim=c(0,10), type="l", main=i,xlab='CSR-R value')
    points(res.natopen.ANO[res.natopen.ANO$fp_ind=="RR1" & res.natopen.ANO$kartleggingsenhet_1m2==i,]$original,
           rep(0,length(res.natopen.ANO[res.natopen.ANO$fp_ind=="RR1" & res.natopen.ANO$kartleggingsenhet_1m2==i,]$original)),
           col="red")
    points(density(res.natopen.ANO[res.natopen.ANO$fp_ind=="RR1" & res.natopen.ANO$kartleggingsenhet_1m2==i,]$original,na.rm=T),
           type="l", col="red")

@@

}
legend("topright", legend=c("reference","field data"), pch=c(NULL,1), lty=1, col=c("black","red"),bty="n", cex=1)

@@

# only two NiN types in GRUK, T2-C-7 and T2-C-8
par(mfrow=c(1,2))
for ( i in unique(res.natopen.GRUK$Kartleggingsenhet) ) {

@@

    plot(density( as.matrix(natopen.ref.cov[['Nitrogen']][,i]) ,na.rm=T),
         xlim=c(1,9), ylim=c(0,2), type="l", main=i,xlab='Nitrogen value')
    points(res.natopen.GRUK[res.natopen.GRUK$fp_ind=="Nitrogen1" & res.natopen.GRUK$Kartleggingsenhet==i,]$original,
           rep(0,length(res.natopen.GRUK[res.natopen.GRUK$fp_ind=="Nitrogen1" & res.natopen.GRUK$Kartleggingsenhet==i,]$original)),
           col="red")
    points(density(res.natopen.GRUK[res.natopen.GRUK$fp_ind=="Nitrogen1" & res.natopen.GRUK$Kartleggingsenhet==i,]$original,na.rm=T),
           type="l", col="red")

@@

}
legend("topright", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"), cex=1)

```
The GRUK figure shows that the distributions for the plant communities' nitrogen affinity in the limestone rich T2-areas (åpen grunnlendt mark) around Oslofjord are shifted towards higher nitrogen affinity. The ANO figure shows shifts towards less ruderal strategy in plant communities limestone poor T2-areas as well as T12 (strandeng) and T16 (rasmarkhei, -eng) types.
